Quantitative Auditing of AI Fairness with Differentially Private Synthetic Data

### 2024-12-09 paper/paper.tex
colors

\documentclass[manuscript,screen,review,anonymous]{acmart}
% \documentclass[acmsmall,sigconf]{acmart}

@@

\begin{table*}[h]
\caption{Fairness measures.}
\label{tab:measures}
\begin{tabular}{llc}
\toprule
\textbf{Category} & \textbf{Fairness Measure} & \textbf{Definition} \\
\midrule
\multirow{4}{*}{Independence} & Disparate Impact & $\frac{P[\hat{Y} = 1 | S \neq 1]}{P[\hat{Y} = 1 | S = 1]} \geq 1 - \epsilon$ \\
& Demographic Parity & $\abs{P[\hat{Y} = 1 | S = 1] - P[\hat{Y} = 1 | S \neq 1]} \leq \epsilon$ \\
& Conditional Statistical Parity & $\abs{P[\hat{Y} = 1 | S = 1, L = l] - P[\hat{Y} = 1 | S \neq 1, L = l]} \leq \epsilon$ \\
& Mean Difference & $\abs{E[\hat{Y}|S = 1] - E[\hat{Y}|S \neq 1]} \leq \epsilon$ \\
\multirow{4}{*}{Separation} & \multirow{2}{*}{Equalized Odds} & $\abs{P[\hat{Y} = 1 | S = 1, Y = 0] - P[\hat{Y} = 1 | S \neq 1, Y = 0]} \leq \epsilon$ \\
& & $\abs{P[\hat{Y} = 1 | S = 1, Y = 1] - P[\hat{Y} = 1 | S \neq 1, Y = 1]} \leq \epsilon$ \\
& Equal Opportunity & $\abs{P[\hat{Y} = 1 | S = 1, Y = 1] - P[\hat{Y} = 1 | S \neq 1, Y = 1]} \leq \epsilon$ \\
& Predictive Equality & $\abs{P[\hat{Y} = 1 | S = 1, Y = 0] - P[\hat{Y} = 1 | S \neq 1, Y = 0]} \leq \epsilon$ \\
\multirow{4}{*}{Sufficiency} & \multirow{2}{*}{Conditional Use Accuracy Equality} & $\abs{P[Y = 1 | S = 1, \hat{Y} = 1] - P[Y = 1 | S \neq 1, \hat{Y} = 1]} \leq \epsilon$ \\
& & $\abs{P[Y = 0 | S = 1, \hat{Y} = 0] - P[Y = 0 | S \neq 1, \hat{Y} = 0]} \leq \epsilon$ \\
& Predictive Parity & $\abs{P[Y = 1 | S = 1, \hat{Y} = 1] - P[Y = 1 | S \neq 1, \hat{Y} = 1]} \leq \epsilon$ \\
& Equal Calibration & $\abs{P[Y = 1 | S = 1, \hat{V} = v] - P[Y = 1 | S \neq 1, \hat{V} = v]} \leq \epsilon$ \\
\multirow{3}{*}{N/A} & Overall Accuracy Equality & $\abs{P[Y = \hat{Y} | S = 1] - P[Y = \hat{Y} | S \neq 1]} \leq \epsilon$ \\
& Positive Balance & $\abs{E[\hat{V} | Y = 1, S = 1] - E[\hat{V} | Y = 1, S \neq 1]} \leq \epsilon$ \\
& Negative Balance & $\abs{E[\hat{V} | Y = 0, S = 1] - E[\hat{V} | Y = 0, S \neq 1]} \leq \epsilon$ \\
\bottomrule
\end{tabular}
\end{table*}

@@

% independence
% \[
% P[\hat{Y} | S] - P[\hat{Y} | \overline{S}]
% = \frac{P[\hat{Y} \cap S]}{P[{S}]} - \frac{P[\hat{Y} \cap \overline{S}]}{P[\overline{S}]}
% \]

% separation
% \[
% P[\hat{Y} | S, Y] - P[\hat{Y} | \overline{S}, Y]
% = \frac{P[\hat{Y} \cap S \cap Y]}{P[{S} \cap Y]} - \frac{P[\hat{Y} \cap \overline{S} \cap Y]}{P[\overline{S} \cap Y]}
% \]

% sufficiency
% \[
% P[Y | S, \hat{Y}] - P[Y | \overline{S}, \hat{Y}]
% = \frac{P[Y \cap S \cap \hat{Y}]}{P[{S} \cap \hat{Y}]} - \frac{P[Y \cap \overline{S} \cap \hat{Y}]}{P[\overline{S} \cap \hat{Y}]}
% \]

>>>>

\documentclass[manuscript,screen,review,anonymous]{acmart}
\usepackage{xcolor}
\usepackage{mathtools}
% \documentclass[acmsmall,sigconf]{acmart}

@@

\begin{table*}[h]
\caption{Fairness measures.}
\label{tab:measures}
\begin{tabular}{llc}
\toprule
\textbf{Category} & \textbf{Fairness Measure} & \textbf{Definition} \\
\midrule
\multirow{4}{*}{Independence} & Disparate Impact & $\frac{P[\hat{Y} = 1 | S \neq 1]}{P[\hat{Y} = 1 | S = 1]} \geq 1 - \epsilon$ \\
& Demographic Parity & $\abs{P[\hat{Y} = 1 | S = 1] - P[\hat{Y} = 1 | S \neq 1]} \leq \epsilon$ \\
& Conditional Statistical Parity & $\abs{P[\hat{Y} = 1 | S = 1, L = l] - P[\hat{Y} = 1 | S \neq 1, L = l]} \leq \epsilon$ \\
& Mean Difference & $\abs{E[\hat{Y}|S = 1] - E[\hat{Y}|S \neq 1]} \leq \epsilon$ \\
\multirow{4}{*}{Separation} & \multirow{2}{*}{Equalized Odds} & $\abs{P[\hat{Y} = 1 | S = 1, Y = 0] - P[\hat{Y} = 1 | S \neq 1, Y = 0]} \leq \epsilon$ \\
& & $\abs{P[\hat{Y} = 1 | S = 1, Y = 1] - P[\hat{Y} = 1 | S \neq 1, Y = 1]} \leq \epsilon$ \\
& Equal Opportunity & $\abs{P[\hat{Y} = 1 | S = 1, Y = 1] - P[\hat{Y} = 1 | S \neq 1, Y = 1]} \leq \epsilon$ \\
& Predictive Equality & $\abs{P[\hat{Y} = 1 | S = 1, Y = 0] - P[\hat{Y} = 1 | S \neq 1, Y = 0]} \leq \epsilon$ \\
\multirow{4}{*}{Sufficiency} & \multirow{2}{*}{Conditional Use Accuracy Equality} & $\abs{P[Y = 1 | S = 1, \hat{Y} = 1] - P[Y = 1 | S \neq 1, \hat{Y} = 1]} \leq \epsilon$ \\
& & $\abs{P[Y = 0 | S = 1, \hat{Y} = 0] - P[Y = 0 | S \neq 1, \hat{Y} = 0]} \leq \epsilon$ \\
& Predictive Parity & $\abs{P[Y = 1 | S = 1, \hat{Y} = 1] - P[Y = 1 | S \neq 1, \hat{Y} = 1]} \leq \epsilon$ \\
& Equal Calibration & $\abs{P[Y = 1 | S = 1, \hat{V} = v] - P[Y = 1 | S \neq 1, \hat{V} = v]} \leq \epsilon$ \\
\multirow{3}{*}{N/A} & Accuracy Equality & $\abs{P[Y = \hat{Y} | S = 1] - P[Y = \hat{Y} | S \neq 1]} \leq \epsilon$ \\
& Positive Balance & $\abs{E[\hat{V} | Y = 1, S = 1] - E[\hat{V} | Y = 1, S \neq 1]} \leq \epsilon$ \\
& Negative Balance & $\abs{E[\hat{V} | Y = 0, S = 1] - E[\hat{V} | Y = 0, S \neq 1]} \leq \epsilon$ \\
\bottomrule
\end{tabular}
\end{table*}

@@

\definecolor{lightgreen}{rgb}{0.6, 1.0, 0.6}
\definecolor{lightred}{rgb}{1.0, 0.6, 0.6}
\newcommand{\mathhighlight}[2]{\colorbox{#1}{$\displaystyle #2$}}
\newcommand{\y}[1]{\mathhighlight{lightgreen}{#1}}
\newcommand{\n}[1]{\mathhighlight{lightred}{#1}}

independence
\[
P[\hat{Y} | S] - P[\hat{Y} | \overline{S}]
= \frac{P[\n{\hat{Y}} \cap \y{S}]}{P[\y{S}]} - \frac{P[\n{\hat{Y}} \cap \y{\overline{S}}]}{P[\y{\overline{S}}]}
\]

separation
\[
P[\hat{Y} | S, Y] - P[\hat{Y} | \overline{S}, Y]
= \frac{P[\n{\hat{Y}} \cap \y{S} \cap \y{Y}]}{P[\y{S} \cap \y{Y}]} - \frac{P[\n{\hat{Y}} \cap \y{\overline{S}} \cap \y{Y}]}{P[\y{\overline{S}} \cap \y{Y}]}
\]

sufficiency
\[
P[Y | S, \hat{Y}] - P[Y | \overline{S}, \hat{Y}]
= \frac{P[\y{Y} \cap \y{S} \cap \n{\hat{Y}}]}{P[\y{S} \cap \n{\hat{Y}}]} - \frac{P[\y{Y} \cap \y{\overline{S}} \cap \n{\hat{Y}}]}{P[\y{\overline{S}} \cap \n{\hat{Y}}]}
\]

accuracy equality
\[
P[Y = \hat{Y} | {S} ] - P[Y = \hat{Y} | \overline{S}]
= \frac{P[\y{Y} = \n{\hat{Y}} \cap \y{S}]}{P[\y{S}]} - \frac{P[\y{Y} = \n{\hat{Y}} \cap \y{\overline{S}}]}{P[\y{\overline{S}}]}
\]
